Supplementary Report for the article *Functional diversity can facilitate the collapse of an undesirable ecosystem state*

### 2023-01-13 reports/supplement/supplement_0.9.1.Rmd
changes to supplement

$H_{j,X}$ can be thought of as a half-inhibition constant: the concentration of $X$ at which the growth rate of species $j$ is reduced by 50%.

@@

Here we provide additional description of the method used to create strain variation within functional groups. The following should be read only after the relevant text in the main report.  We generated variation in maximum growth rate $G_{max}$ and tolerance to reduced sulfur $h_{SR}$ or oxygen $h_{O}$ by varying the range of trait values around a reference trait value. To manipulate the amount of variation in a given trait and functional group, we introduced the meta-parameter $Div$ subscripted by the trait and the functional group. Take for example the cyanobacteria functional group, a reference maximum growth rate parameter $G_{max,CB}$ and a reference tolerance parameter $h_{SR,CB}$. The amount of among strain variation in the cyanobacteria is controlled by the two meta-parameters $Div_{G_{max,CB}}$ and $Div_{h_{SR,CB}}$. To create a trade-off, the two meta-parameters were always of different sign. The among strain variation was calculated such that the growth rate of strain $i = 1$ was a factor $1/({2Div_{G_{max,CB}}})$ of the reference growth rate, and the growth rate of the $i = n$ strain was a factor $2Div_{G_{max,CB}}$ of the reference growth rate. The growth rate of the $i = {2,...n-1}$ other strains was the reference growth rate multiplied by a factor that was equally distributed between the factor of the $i = 1$ and $i = n$ strain. This implementation of the variation and trade-off has the assumption of a linear trade-off of log-log transformed among-strain variation. Hence, for example, with the reference growth rate $G_{max,CB} = 0.05$, among strain variation of $Div_{G_{max,CB}} = 1$, and nine strains ($n = 9$), the growth rates of the nine strains are 0.025, 0.03, 0.035, 0.042, 0.05, 0.059, 0.071, 0.084, 0.1. Put another way, the growth rate parameter of the $i$-th of the $n$ strains in a functional group was calculated as $G_{max,i} = D_i * G_{max}$ where $D_i$ is the $i$-th element of the set $D = {2f(x)Div_{G_{max}} |x ∈ N, x ≤ n}$, where $f(x) = (x − (n + 1)/2)/((n − 1)/2)$.

@@

All three examples have the following amounts of trait variation: $Div_{G_{max,CB}}$ = `r round(CB_gmax_div, 3)`, $Div_{h_{SR,CB}}$ = `r round(CB_h_div, 3)`, $Div_{G_{max,SB}}$ = `r round(SB_gmax_div, 3)`, $Div_{h_{O,SB}}$ = `r round(SB_h_div, 3)`, $Div_{G_{max,PB}}$ = `r round(PB_gmax_div, 3)`, $Div_{h_{O,PB}}$ = `r round(PB_h_div, 3)`. These are also the maximum amounts of trait variation explored in the simulations.

>>>>

$H_{j,X}$ can be thought of as a half-inhibition constant: the concentration of $X$ at which the growth rate of species $j$ is reduced by 50%. A higher half-inhibition constant therefore means higher tolerance of species $j$ to substrate $X$.

@@

Here we provide additional description of the method used to create strain variation within functional groups. The following should be read only after the relevant text in the main report.  We generated variation in maximum growth rate $g_{max}$ and in tolerance to reduced sulfur or oxygen (i.e. in the half-inhibition constants $H_{SR}$ or $H_{O}$) by varying the range of trait values around a reference trait value. To manipulate the amount of variation in a given trait and functional group, we introduced the meta-parameter $Div$ subscripted by the trait and the functional group. Take for example the cyanobacteria functional group, a reference maximum growth rate parameter $g_{max,CB}$ and a reference tolerance parameter $H_{SR,CB}$. The amount of among strain variation in the cyanobacteria is controlled by the two meta-parameters $Div_{g_{max,CB}}$ and $Div_{H_{SR,CB}}$. To create a trade-off, the two meta-parameters were always of different sign. The among strain variation was calculated such that the growth rate of strain $i = 1$ was a factor $1/({2Div_{g_{max,CB}}})$ of the reference growth rate, and the growth rate of the $i = n$ strain was a factor $2Div_{g_{max,CB}}$ of the reference growth rate. The growth rate of the $i = {2,...n-1}$ other strains was the reference growth rate multiplied by a factor that was equally distributed between the factor of the $i = 1$ and $i = n$ strain. This implementation of the variation and trade-off has the assumption of a linear trade-off of log-log transformed among-strain variation. Hence, for example, with the reference growth rate $g_{max,CB} = 0.05$, among strain variation of $Div_{g_{max,CB}} = 1$, and nine strains ($n = 9$), the growth rates of the nine strains are 0.025, 0.03, 0.035, 0.042, 0.05, 0.059, 0.071, 0.084, 0.1. Put another way, the growth rate parameter of the $i$-th of the $n$ strains in a functional group was calculated as $g_{max,i} = D_i * g_{max}$ where $D_i$ is the $i$-th element of the set $D = {2f(x)Div_{g_{max}} |x ∈ N, x ≤ n}$, where $f(x) = (x − (n + 1)/2)/((n − 1)/2)$.

@@

All three examples have the following amounts of trait variation: $Div_{g_{max,CB}}$ = `r round(CB_gmax_div, 3)`, $Div_{H_{SR,CB}}$ = `r round(CB_h_div, 3)`, $Div_{g_{max,SB}}$ = `r round(SB_gmax_div, 3)`, $Div_{H_{O,SB}}$ = `r round(SB_h_div, 3)`, $Div_{g_{max,PB}}$ = `r round(PB_gmax_div, 3)`, $Div_{H_{O,PB}}$ = `r round(PB_h_div, 3)`. These are also the maximum amounts of trait variation explored in the simulations.
